---
title: "03-visualise-scot"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(tidytext)
library(lubridate)
library(here)
library(scales)
library(glue)

# set theme for plots to minimal
theme_set(theme_minimal(base_size = 12))

# set color
# https://www.schemecolor.com/flag-of-scotland-colors.php
scotblue <- "#0065BF"

# figure sizing
knitr::opts_chunk$set(
  fit.width = 6,
  fig.asp = 0.618,
  fig.retina = 3, 
  out.width = "100%",
  dpi = 300
)

# read data
covid_speeches_scot <- read_rds(here::here("processed-data", "covid-speeches-scot.rds"))
covid_speeches_scot_bigrams <- read_rds(here::here("processed-data", "covid-speeches-scot-bigrams.rds"))
covid_speeches_scot_words <- read_rds(here::here("processed-data", "covid-speeches-scot-words.rds"))
```

## Length of speech

```{r}
ggplot(covid_speeches_scot, aes(x = n_words)) +
  geom_density(color = scotblue, fill = scotblue, alpha = 0.5) +
  labs(
    title = "Distribution of number of words",
    subtitle = "of Scotland daily briefings",
    x = "Number of words",
    y = "Density"
  ) +
  theme(axis.text.y = element_blank())
```

```{r}
lm_words <- lm(n_words ~ date, data = covid_speeches_scot)
lm_words_rsq <- glance(lm_words)$r.squared

covid_speeches_scot %>%
  ggplot(aes(x = date, y = n_words)) +
  geom_point(color = scotblue, alpha = 0.7) +
  geom_smooth(aes(x = date, y = n_words), method = lm, formula = y ~ x, color = "darkgray") +
  labs(
    title = "Length of Scotland COVID-19 speeches",
    subtitle = glue("Measured in number of words, R-squared = {percent(lm_words_rsq)}"),
    x = NULL, y = "Number of words", color = NULL, shape = NULL
  )
```

## Word frequency

```{r fig.asp = 0.8}
covid_speeches_scot_words %>%
  count(word, sort = TRUE) %>%
  filter(n > 1000) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n, fill = n)) +
  geom_col() +
  guides(fill = FALSE)
```

## Remove stop words

```{r remove-stop-words}
covid_speeches_scot_words <- covid_speeches_scot_words %>%
  anti_join(stop_words)
```

## Word frequency again

```{r fig.asp = 0.8}
threshold <- 500

covid_speeches_scot_words %>%
  count(word, sort = TRUE) %>%
  filter(n > threshold) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n, fill = log(n))) +
  geom_col(show.legend = FALSE) +
  labs(
    title = "Frequency of words in Scotland COVID-19 briefings",
    subtitle = glue("Words occurring more than {threshold} times"),
    y = NULL, x = NULL
  )
```

## Sentiment analysis

Sentiments: Positive and negative.

```{r fig.asp = 0.8}
covid_speeches_scot_words %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(sentiment, word, sort = TRUE) %>%
  group_by(sentiment) %>%
  slice_head(n = 20) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n, fill = sentiment)) +
  geom_col() +
  guides(fill = FALSE) +
  facet_wrap(~ sentiment, scales = "free") +
  labs(
    title = "Sentiment and frequency of words in Scotland COVID-19 briefings",
    subtitle = "Bing lexicon",
    y = NULL, x = NULL
  )
```

"positive" isn't really a positive word in this context. Remove and plot again.

```{r fig.asp = 0.8}
covid_speeches_scot_words %>%
  filter(word != "positive") %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(sentiment, word, sort = TRUE) %>%
  group_by(sentiment) %>%
  slice_head(n = 20) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n, fill = sentiment)) +
  geom_col() +
  guides(fill = FALSE) +
  facet_wrap(~ sentiment, scales = "free") +
  labs(
    title = "Sentiment and frequency of words in Scotland COVID-19 briefings",
    subtitle = "Bing lexicon",
    y = NULL, x = NULL
  )
```

## Daily sentiments

### Lexicon: Bing

Sentiments: Positive and negative.

```{r}
covid_speeches_scot_words %>%
  filter(word != "positive") %>% 
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(date, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n) %>%
  mutate(sentiment = positive - negative) %>%
  ggplot(aes(x = date, y = sentiment)) +
  geom_line(color = "gray") +
  geom_point(aes(color = sentiment > 0, shape = sentiment > 0), size = 2, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "lightgray") +
  labs(
    title = "Daily sentiment score of words in Scotland COVID-19 briefings",
    subtitle = "Bing lexicon",
    x = "Date", y = "Sentiment score (positive - negative)"
  )
```

and now with a smooth curve...

```{r}
covid_speeches_scot_words %>%
  filter(word != "positive") %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(date, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n) %>%
  mutate(sentiment = positive - negative) %>%
  ggplot(aes(x = date, y = sentiment)) +
  geom_smooth(color = "gray", method = "loess", formula = y ~ x) +
  geom_point(aes(color = sentiment > 0, shape = sentiment > 0), size = 1, show.legend = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "lightgray") +
  labs(
    title = "Sentiment score Scotland COVID-19 briefings",
    subtitle = "Sentiment score is calculated as number of positive - number of negative words, according to the Bing lexicon",
    x = "Date", y = "Sentiment score (positive - negative)"
  )
```

### Lexicon: NRC

Sentiments: trust, fear, negative, sadness, anger, surprise, positive, disgust, joy, and anticipation.

```{r fig.asp=0.7}
covid_speeches_scot_words %>%
  filter(word != "positive") %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  mutate(
    sentiment = fct_relevel(sentiment, "positive", "anticipation", "joy", "surprise", "trust",
                            "negative", "anger", "disgust", "fear", "sadness"),
    sentiment_binary = if_else(sentiment %in% c("positive", "anticipation", "joy", "surprise", "trust"), "positive", "negative")
  ) %>%
  count(sentiment_binary, sentiment, word, sort = TRUE) %>%
  group_by(sentiment) %>%
  slice_head(n = 10) %>%
  ggplot(aes(y = fct_reorder(word, n), x = n, fill = sentiment_binary)) +
  geom_col() +
  guides(fill = FALSE) +
  facet_wrap(~ sentiment, scales = "free_y", ncol = 5) +
  labs(
    title = "Sentiment and frequency of words in Scotland COVID-19 briefings",
    subtitle = "NRC lexicon",
    y = NULL, x = NULL
  ) +
  scale_x_continuous(breaks = c(0, 1000)) +
  theme_minimal(base_size = 11)
```

```{r}
covid_speeches_scot_words %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  mutate(
    sentiment = fct_relevel(sentiment, "positive", "anticipation", "joy", "surprise", "trust",
                            "negative", "anger", "sadness", "disgust", "fear"),
    sentiment_binary = if_else(sentiment %in% c("positive", "anticipation", "joy", "surprise", "trust"), "positive", "negative")
    ) %>%
  count(date, sentiment_binary, sentiment) %>%
  ggplot(aes(x = date, y = n, color = sentiment_binary)) +
  geom_line(size = 0.3) +
  guides(color = FALSE) +
  facet_wrap(~ sentiment, ncol = 5) +
  labs(
    title = "Sentiment score of words in Scotland COVID-19 briefings over time",
    subtitle = "NRC lexicon",
    x = "Date", y = "Sentiment score", color = NULL
  ) +
  scale_x_date(guide = guide_axis(check.overlap = TRUE))
```

## Bigram frequency

```{r fig.asp=0.8}
threshold <- 75

covid_speeches_scot_bigrams %>%
  mutate(
    bigram = if_else(bigram == "care home", "care home(s)", bigram),
    bigram = if_else(bigram == "care homes", "care home(s)", bigram)
  ) %>%
  count(bigram, sort = TRUE) %>%
  filter(n > threshold) %>%
  ggplot(aes(y = fct_reorder(bigram, n), x = n, fill = n)) +
  geom_col() +
  guides(fill = FALSE) +
  labs(
    title = "Frequency of bigrams in Scotland COVID-19 briefings",
    subtitle = glue("Bigrams occurring more than {threshold} times"),
    y = NULL, x = NULL
  )
```

## Social vs. physical distancing

```{r}
covid_speeches_scot %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  filter(str_detect(bigram, "social dist|physical dist")) %>%
  mutate(soc_phys = if_else(str_detect(bigram, "social"), "S", "P")) %>%
  count(date, soc_phys) %>%
  ggplot(aes(x = date, y = n, color = soc_phys)) +
  geom_text(aes(label = soc_phys)) +
  guides(color = FALSE) +
  labs(x = "Date", y = "Frequency",
       title = "Social (S) vs. physical (P) distancing",
       subtitle = "Number of mentions over time") +
  scale_color_manual(values = c(scotblue, "darkgray")) +
  scale_y_continuous(limits = c(0, 10), breaks = seq(0, 10, 2))
```

## Vaccines

```{r dev="ragg_png"}
covid_speeches_scot_words %>%
  filter(str_detect(word, "[Vv]accin")) %>%
  count(date) %>%
  ggplot(aes(x = date, y = n)) +
  geom_text(aes(label = "💉", size = n), show.legend = FALSE) +
  labs(x = "Date", y = "Frequency",
       title = 'Number of times "vaccine" is mentioned in speech')
```

