---
title: "08-visualise-compare"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(tidytext)
library(lubridate)
library(here)
library(scales)
library(glue)

# set theme for plots to minimal
theme_set(theme_minimal(base_size = 12))

# set color
scotblue <- "#0065BF"
ukred <- "#D00C27"

# figure sizing
knitr::opts_chunk$set(
  fit.width = 6,
  fig.asp = 0.618,
  fig.retina = 3, 
  out.width = "100%",
  dpi = 300
)

# read data
covid_speeches_words <- read_rds(here::here("processed-data", "covid-speeches-words.rds"))
```

## Basic comparison

More speeches in Scotland than UK and longer on average.

```{r}
covid_speeches_words %>%
  group_by(origin) %>%
  summarise(
    n_speeches = max(speech_no),
    n_words    = n(),
    avg_words  = n_words / n_speeches,
    .groups = "drop"
    )
```

```{r fig.asp = 0.5}
covid_speeches_words %>%
  distinct(origin, speech_no, .keep_all = TRUE) %>%
  ggplot(aes(x = n_words, color = origin, fill = origin)) +
  geom_density(alpha = 0.7) +
  scale_color_manual(values = c(scotblue, ukred)) +
  scale_fill_manual(values = c(scotblue, ukred)) +
  labs(
    x = "Number of words",
    y = "Density",
    color = "Origin", fill = "Origin"
  ) +
  theme(axis.text.y = element_blank())
```

## TF-IDF

The statistic tf-idf is intended to measure how important a word is to a document in a collection (or corpus) of documents, for example, to one novel in a collection of novels or to one website in a collection of websites.^[Source: [Text Mining with R](https://www.tidytextmining.com/tfidf.html)]

Calculate tf-idf

```{r}
covid_speeches_tf_ifd <- covid_speeches_words %>%
  count(origin, word, sort = TRUE) %>%
  group_by(origin) %>%
  mutate(total = sum(n)) %>%
  bind_tf_idf(word, origin, n)
```

View words with high TF-IDF

```{r}
covid_speeches_tf_ifd %>%
  select(-total) %>%
  arrange(desc(tf_idf))
```

Visualise TF-IDF

```{r}
covid_speeches_tf_ifd %>%
  arrange(desc(tf_idf)) %>%
  mutate(word = factor(word, levels = rev(unique(word)))) %>% 
  group_by(origin) %>% 
  slice_head(n = 15) %>% 
  ungroup() %>%
  ggplot(aes(y = word, x = tf_idf, fill = origin)) +
  geom_col(show.legend = FALSE) +
  labs(y = NULL, x = "tf-idf", title = "Common words in COVID briefings") +
  facet_wrap(~origin, ncol = 2, scales = "free") +
  scale_fill_manual(values = c(scotblue, ukred)) +
  scale_x_continuous(breaks = c(0, 0.00015, 0.0003), labels = label_number())
```

Some differences observed:

- "reclaiming our freedoms", "reclaim our lives in full" etc. in UK speeches 